---
title: "Intro to GIS using R"
output: html_document
date: "2023-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# load packages
library(sf) # for geospatial analysis
library(tidyverse) # for all things data wrangling
library(tigris) # for downloading administrative boundary shapefiles
library(ggthemes) # for the extra map theme
library(here) # use to make file paths much easier and more reproducible
library(measurements) # use for converting measurements
library(leaflet) # for an interactive map
library(htmlwidgets) # for saving out interactive map
```

## Learn how to use basic sf functions and create both static and interactive maps!


### **Data Sources:**

- [Sonoma County Schools](https://gis-sonomacounty.hub.arcgis.com/datasets/sonomacounty::schools-1/about)
- [Sonoma County Regional Parks](https://gis-sonomacounty.hub.arcgis.com/datasets/sonomacounty::regional-parks-boundaries/about)


### Read in our data:

```{r, echo=FALSE}
# Path to raw data folder
raw_data <- here("intro_gis_r/data/raw/")

# read in our shapefiles
schools <- sf::read_sf(file.path(raw_data, "soco_schools/FAC_SCHL.shp"))
parks <- sf::read_sf(file.path(raw_data, "soco_parks/BND_PRK.shp"))

# Get administrative boundaries using the tigris package
  # Never search for a state/county/census boundary shapefile online ever again!!
us_states <- tigris::states(cb = TRUE) # cb = TRUE downloads the 500k resolution version
ca_counties <- tigris::counties(state = "CA", cb = TRUE)
```

### Look at our data:

```{r, echo=FALSE}
# attribute data
view(us_states)
```

```{r}
# Map US states and CA counties
ggplot() +
  geom_sf(data = us_states) +
 # geom_sf(data = ca_counties, color = "red") +
 # geom_sf(data = ca_counties, aes(color = NAME), show.legend = FALSE) + # color the outlines of counties by their name
 geom_sf(data = ca_counties, aes(fill = ALAND)) + # color the fill of counties by the area of land they cover
  coord_sf(xlim = c(-125, -65), ylim = c(23,50)) +
  theme_map()
```

```{r, echo=FALSE}
view(schools)
```

```{r}
# Map location of schools in Sonoma County
ggplot() +
  geom_sf(data = ca_counties) +
#  geom_sf(data = schools, color = "red") +
  geom_sf(data = schools, aes(color = PublicSchl)) +
  coord_sf(xlim = c(-123.5,-122.2), ylim = c(38,39)) +
   labs(
     color = "Public School"
   ) +
  theme_map()
```

```{r}
# Let's make a quick basemap
soco_basemap <- ggplot() +
  geom_sf(data = ca_counties, fill = "grey95") +
  geom_sf(data = ca_counties %>% filter(NAME == "Sonoma"), fill = "grey85")

soco_basemap +
  coord_sf(xlim = c(-123.5,-122.2), ylim = c(38,39)) +
  theme_map()
```


```{r, echo=FALSE}
view(parks)
```

```{r}
# Map location of parks in Sonoma County
soco_basemap +
  geom_sf(data = parks, fill = "green", color = "darkgreen") +
  coord_sf(xlim = c(-123.5,-122.2), ylim = c(38,39)) +
  theme_map()

```

## TASK: Find out which scools are within a 1 mi radius of regional parks

1. Make sure all Coordinate Reference Systems (CRS) match.

```{r}
# Let's check what the CRS looks like. Note the EPSG code, and the unit of measurement
st_crs(parks)
```

```{r}
# now let's see if they match
st_crs(parks) == st_crs(schools)
```
```{r}
# Those match but let's transform the county shapefile to match
ca_counties_crs <- sf::st_transform(ca_counties, st_crs(parks))

# check if it worked
st_crs(parks) == st_crs(ca_counties_crs)
```


2. Buffer the parks shapefile by 1 mi

```{r, echo=FALSE}
# get the number of ft in a mile
ft_1mi <- measurements::conv_unit(1, "mi", "ft")
  
# buffer
parks_buffer <- st_buffer(parks, ft_1mi)
```

```{r}
# check it worked
soco_basemap +
  geom_sf(data = parks_buffer, fill = "blue") + # 1 mile buffer
  geom_sf(data = parks, color = "green", fill = "green") + # original parks shapefile 
  coord_sf(xlim = c(-123.5,-122.2), ylim = c(38,39)) +
  theme_map()
```

3. Intersect the buffer with schools

```{r}
schools_1mi_parks <- st_intersection(schools, parks_buffer)
```

```{r}
# check this worked
soco_basemap +
  geom_sf(data = parks_buffer, fill = "blue") +
  geom_sf(data = schools_1mi_parks, color = "yellow") +
  coord_sf(xlim = c(-123.5,-122.2), ylim = c(38,39)) +
  theme_map()
```

5. Map our results

```{r}
final_map <- soco_basemap +
 # geom_sf(data = parks, fill = "dark green") +
  geom_sf(data = schools_1mi_parks, aes(fill = PublicSchl), shape = 24, size = 2) +
  scale_fill_manual(values = c("orange2", "purple")) +
  labs(
    color = "Public School"
  ) +
  coord_sf(xlim = c(-123.5,-122.3), ylim = c(38.13,38.85)) +
  theme_map()

final_map
```

```{r, include = FALSE}
# # save our map
 ggsave(filename = "schls_parks.png", plot = final_map, 
        path = "C:/Users/PMcKibben/OneDrive - Eastern Research Group/github/erg_data_science_training_acforks", dpi = 350)
# line 183, code often doesn't like the "here" function
```

Optional - Functionize!

```{r}
# Create a function to easily replicate our workflow above
process_shps = function(shp_int, shp_buff, buffer_dist){
  
  # Make sure the CRS matches <-THIS IS VERY IMPORTANT IN ANY GIS ANALYSIS
  shp_int <- st_transform(shp_int, st_crs(shp_buff))
  
  # Buffer 
  shp_buff <- st_buffer(shp_buff, buffer_dist)
  
  # Intersect
  shp_final <- st_intersection(shp_int, shp_buff)
  
  return(shp_final)
}
```

```{r}
schls_parks_6000ft <- process_shps(schools, parks, 6000) #schools main thing we're intercepting, parks is the buffer, 6000 is the ft buffered
```

Optional - Interactive Map using Leaflet 

Check out their [documentation](https://rstudio.github.io/leaflet/)

```{r}
# Make a quick interactive map
m <- leaflet(data = schls_parks_6000ft) %>% 
 # addTiles() %>% if you only use the code in this line, it will use leaflet's base map
  addProviderTiles(providers$CartoDB.Positron) %>% #adds a diferent base map
  addMarkers(~Long, ~Lat, popup = ~School) #controls what you see when you click on markers

m
```

```{r, include = FALSE}
# let's save the map to share with coworkers
 saveWidget(m, file="schls_interactive.html")
```

